\LARGE \bf{Hall C 12 GeV Trigger Set-Up}

### 2020-08-30 MY_THESIS/HallC_12GeV_Trigger/version3/trigger_v3.tex
updated trigger doc

\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}
\usepackage[title]{appendix}
\usepackage[section]{placeins}
\usepackage{float}

@@

@@

\begin{figure}[h!]
  \includegraphics[scale=0.5]{../CR_patch_combined.png}
  \caption{(A) Counting Room Patch Panels (left 2 racks). (B) Counting Room main
    Electronic Racks for HMS/SHMS detectors (right 2 racks).}
  \label{fig:cr_racks_patch}
\end{figure}
\newpage
\section{HMS Trigger Set-Up} 
\indent The XY scintillator arrays (hodoscope planes) will form part of the standard HMS trigger configuration.  
Additional particle detectors may also be incorporated into the HMS trigger as required by different experiments. The Gas \v{C}erenkov and Calorimeter triggers will be
used for e/$\pi$ separation, whereas the Aerogel \v{C}erenkov trigger will be used for $\pi$/K/p separation.

\subsection{Hodoscope Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hHODO_diagram.png}
  \caption{HMS Hodoscope Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA1}}
  \label{fig:hHODO_diagram}
\end{figure}
\noindent Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:hms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:hHODO_diagram},
for example, hodoscope plane h1X consists of 32 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x16 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC). The remaining two-thirds (66\%)
of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The HMS discriminators thresholds and gate widths were set to \hhodthrs and \hhodgate, respectively.\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to the pin corresponding to the
desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{h1X = h1X+ (16-fold OR) \textit{AND} h1X- (16-fold OR)} \\ 
& \text{h1Y = h1Y+ (10-fold OR) \textit{AND} h1Y- (10-fold OR)} \\
& \text{h2X = h2X+ (16-fold OR) \textit{AND} h2X- (16-fold OR)} \\ 
& \text{h2Y = h2Y+ (10-fold OR) \textit{AND} h2Y- (10-fold OR)} 
\end{empheq}

@@

\subsection{Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hCAL_diagram.png}
  \caption{HMS Calorimeter Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA2}}
  \label{fig:hCAL_diagram}
\end{figure}
\noindent The HMS Calorimeter consists of four layers of lead blocks. Layers A and B read out a 26 PMT signals per layer (13 signals/side) while layers C and D
read out 13 signals/layer on one side. The first two layers form the Pre-Shower Counter while all four layers (A, B, C and D) form the Shower Counter.  
Each layer is read out in the Counting Room patch and fed into 50:50 splitters. One output of the splitter is fed to fADCs via a Ribbon-to-BNC converter (same as hodoscopes)
while the other output is sent to a P/S Model 740 NIM Linear FI/FO summing modules. Each side of a layer is summed first (hA+, hA-, hB+, hB-, hC and hD sums). The sums are fed into
a LeCroy Model 428F summing module where layers hA+/- and hB+/- are summed to form hA and hB sums. A copy of each layer sum is sent to fADCs. The Pre-Shower SUM is then made from the sum of layers A and B, while the Shower SUM is made by summing all four layers. A copy of the PreShower and Shower sums is also sent to fADCs. The Pre-Shower and Shower sums are then sent to a P/S Model 715
NIM Discriminator unit to form the PreShower Low/High (LO/HI) and Shower Low pre-triggers with thresholds \hPrShLo, \hPrShHi and \hSHLo, respectively with all gate widths set to 30 ns. A copy of the pre-triggers is sent to TDCs/Scaler modules for trigger timing and counting rate
information.

\subsection{Gas \v{C}erenkov Pre-Trigger} \label{hms_cer_section}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{hCER_diagram.png}
  \caption{HMS Gas \v{C}erenkov Electronics Diagram. Same electronics diagram applies for SHMS Gas \v{C}erenkov. For original diagram, see Appendix \ref{appendix:Appx3} }
  \label{fig:hCER_diagram}
\end{figure}

@@

\newpage
\subsection{HMS Single Arm Pre-Trigger}\label{ssec:hms_single_arm_sec}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{HMS_SingleArm_diagram.png}
  \caption{HMS Single Arm Pre-Trigger Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx5}}
  \label{fig:HMS_SingleArm_diagram}
\end{figure}

@@

\subsection{Hodoscopes Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pHODO_diagram.png}
  \caption{SHMS Hodoscopes Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx6}}
  \label{fig:pHODO_diagram}
\end{figure}

Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:shms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:shms_hod_trg},
for example, hodoscope plane S1X consists of 26 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x13 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC).
The remaining two-thirds (66\%) of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The SHMS discriminators thresholds and gate widths were set to \shodthrs and \shodgate, respectively,
whereas the Quartz Discriminators thresholds and gate widths were set to \quartzthrs and \shodgate, respectively\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to
the pin corresponding to the desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{S1X = S1XL (13-fold OR) \textit{AND} S1XR (13-fold OR)} \\ 
& \text{S1Y = S1YT (14-fold OR) \textit{AND} S1YB (14-fold OR)} \\
& \text{S2X = S2XL (13-fold OR) \textit{AND} S2XR (13-fold OR)} \\ 
  & \text{S2Y = \big\{S2Y[1-16]T OR S2Y[17-21]T\big\} \textit{AND} \big\{S2Y[1-16]B OR S2Y[17-21]B\big\}}
\end{empheq}

\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal
(twisted pair) to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence.
An X-Y hodoscope plane coincidence (S1 = S1X \textit{AND} S1Y, S2 = S2X \textit{AND} S2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four
individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or 4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the
pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:pHODO_diagram}) \\

\subsection{Pre-Shower / Shower Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pCAL_diagram.png}
  \caption{SHMS PreShower / Shower Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx7}}
  \label{fig:pCAL_diagram}
\end{figure}

The SHMS Pre-Shower consists of two sets of fourteen (PMT-coupled) lead blocks oriented perpendicular to the Shower Counter blocks\cite{shms_preSh_talk}. The initial sum was done in the SHMS Electronics Hut. PMT signals
from groups of four blocks were summed to form:
\begin{empheq}[box=\fbox]{align*}
  & \text{preSh SUM [1-4]: [1-4]L + [1-4]R} \\ 
  & \text{preSh SUM [5-8]: [5-8]L + [5-8]R} \\ 
  & \text{preSh SUM [9-12]: [9-12]L + [9-12]R} \\ 
  & \text{preSh SUM [13-14]: [13-14]L + [13-14]R}
\end{empheq}
The Shower counter consists of 224 lead blocks, each coupled to a PMT at the end. Becasue of the high channel density of this detector, its signals are sent
directly to the ROC 4 fADCs in the SHMS detector hut.
\newpage
\subsection{Heavy/Noble Gas \v{C}erenkov Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{pCER_diagram.png}
  \caption{SHMS Gas \v{C}erenkovs Electronics Diagram.}
  \label{fig:pCER_diagram}
\end{figure}
\noindent The SHMS Heavy Gas \v{C}erenkov detector consists of a 1 m long, 1.6 m in diameter cylindrical tank located between the first set of hodoscope planes and the Aerogel (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four thin spherical mirrors that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_hgc_talk}. 
\newline
\indent The SHMS Noble Gas \v{C}erenkov detector consists of a 2 m long active length of Argon/Neon gas tank located before the first drift chamber (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four overlapping mirrors, that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_ngc_talk}. 
For the original electronics diagram, See Appendix \ref{appendix:Appx4} on the HMS Gas \v{C}erenkov.
\subsection{Aerogel \v{C}erenkov Pre-Trigger}
\indent The SHMS Aerogel \v{C}erenkov detector consists of a 110 x 100 x 24.5 cm$^{3}$ rectangular aerogel tray coupled to a diffusion box. The diffusion box has seven 5-inch PMTs on each side
which detect \v{C}erenkov light produced by interactions with the Aerogel material\cite{shms_aero_article}. The detector is located between Heavy Gas \v{C}erenkov and second set of hodoscope planes
(See Figure \ref{fig:shms_stack}). The electronics diagram is the same as in Figure \ref{fig:pCER_diagram}. For the original diagram, see Appendix \ref{appendix:Appx4}
\newpage
\subsection{SHMS Single Arm Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{SHMS_SingleArm_diagram.png}
  \caption{SHMS Single Arm Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx8}.}
  \label{fig:SHMS_SingleArm_diagram}
\end{figure}

\indent The SHMS single arm trigger is be formed exactly as the HMS single arm trigger, with the exception of the detectors involved. Compare the electronics diagrams in Figures \ref{fig:SHMS_SingleArm_diagram} and
\ref{fig:HMS_SingleArm_diagram}, and read Section \ref{ssec:hms_single_arm_sec} for a detailed description of the electronic diagrams. 
\newpage
\section{Coincidence Trigger Set-Up}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.5]{Coin_diagram.png}
  \caption{Coincidence Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx9}}
  \label{fig:Coin_TRG}
\end{figure}
\indent In coincidence mode, the HMS and SHMS pre-triggers are sent to a NIM Logic module, where the first pre-trigger that arrives will open
a coincidence time window during which the second pre-trigger may or may not arrive in time. This will determine whether two events that occurred in
each spectrometer are correlated with the event that originated at the target. If the coincidence pre-trigger is formed, a copy will be sent to
scalers and TDCs, while the other copy is sent to a NIM/ECL converter where three copies (may be more) are sent via twisted pairs to every TDC module and the TI, which will be the \textbf{ONLY} Trigger Master in coincidence mode. Additional copies (not shown) may need to be sent to the HMS readout crate (ROC 01) as well, since the coincidence pre-trigger\footnote{A copy of the coincidence pre-trigger will also need to be sent to the
  HMS/SHMS TDC modules present in the detector huts crates via the patch panel.} is common to all crates. \\
\indent Once the coincidence pre-trigger is processed by the TI Master in ROC 02, it is sent to a NIM/ECL converter, where a copy is sent to scalers and TDCs, and the other copy is sent to the ``TRG'' NIM input in the TDC front panel, and daisy-chained with other TDCs in ROC02 and ROC01. The remaining crates in the HMS/SHMS huts receive a copy of the accepted trigger via fiber optics lines running from the TI Master to the TIs in all other crates. The TIs then distribute the accepted trigger to all modules of their respective crates. 
\newpage
\onecolumn
\bibliography{trigger_v2}
\bibliographystyle{acm}

>>>>

\usepackage{url}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{graphicx}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage[top=0.6 in, bottom=0.7 in, left=0.7in, right=0.7in]{geometry}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{vector}
\usepackage{amsmath}
\usepackage{empheq}
\usepackage{amssymb}
\usepackage{tgtermes}
\usepackage{gensymb}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{epstopdf}
\usepackage[singlelinecheck=false]{caption}
\usepackage[hyperfootnotes=false, pdfnewwindow, colorlinks=true]{hyperref}
\usepackage[title]{appendix}
\usepackage[section]{placeins}
\usepackage{float}
\usepackage{xcolor}

@@

\begin{center}
  \vspace*{\stretch{1}}
  \textbf{\noindent \large \textcolor{red}{DISCLOSURE: \\ THE ELECTRONIC DIAGRAMS SHOWN IN THE MAIN TEXT ARE UPDATED AND CORRESPOND TO THE ELECTRONICS DIAGRAMS USED IN A1N/D2N EXPERIMENTS AS OF FALL 2019 - FALL 2020. THE CHANGES ARE MINOR, HOWEVER, IF YOUR EXPERIMENT RAN BEFORE FALL 2019, PLEASE REFER TO SECTION 6 (PAGE 17) OF THIS DOCUMENT FOR CLARIFICATION OF THE CHANGES MADE.}}
  \vspace*{\stretch{1}}
\end{center}
\newpage

@@

\begin{figure}[h!]
  \includegraphics[scale=0.5]{../CR_patch_combined.png}
  \caption{(A) Counting Room Patch Panels (left 2 racks). (B) Counting Room main
    Electronic Racks for HMS/SHMS detectors (right 2 racks).}
  \label{fig:cr_racks_patch}
\end{figure}

\newpage
\section{HMS Trigger Set-Up} 
\indent The XY scintillator arrays (hodoscope planes) will form part of the standard HMS trigger configuration.  
Additional particle detectors may also be incorporated into the HMS trigger as required by different experiments. The Gas \v{C}erenkov and Calorimeter triggers will be
used for $e/\pi$ separation, whereas the Aerogel \v{C}erenkov trigger will be used for $\pi/K/p$ separation.

\subsection{Hodoscope Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{./hHODO_diagram.png}
  \caption{HMS Hodoscope Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA1}}
  \label{fig:hHODO_diagram}
\end{figure}
\noindent Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:hms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:hHODO_diagram},
for example, hodoscope plane h1X consists of 32 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x16 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC). The remaining two-thirds (66\%)
of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The HMS discriminators thresholds and gate widths were set to \hhodthrs and \hhodgate, respectively.\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to the pin corresponding to the
desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{h1X = h1X+ (16-fold OR) \textit{AND} h1X- (16-fold OR)} \\ 
& \text{h1Y = h1Y+ (10-fold OR) \textit{AND} h1Y- (10-fold OR)} \\
& \text{h2X = h2X+ (16-fold OR) \textit{AND} h2X- (16-fold OR)} \\ 
& \text{h2Y = h2Y+ (10-fold OR) \textit{AND} h2Y- (10-fold OR)} 
\end{empheq}

@@

\subsection{Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{./hCAL_diagram.png}
  \caption{HMS Calorimeter Electronics Diagram. For original diagram, see Appendix \ref{appendix:AppxA2}}
  \label{fig:hCAL_diagram}
\end{figure}
\noindent The HMS Calorimeter consists of four layers of lead blocks. Layers A and B read out a 26 PMT signals per layer (13 signals/side) while layers C and D
read out 13 signals/layer on one side. The first layer form the Pre-Shower counter while all four layers (A, B, C and D) form the Shower counter.  
Each layer is read out in the Counting Room patch and fed into 50:50 splitters. One output of the splitter is fed to fADCs via a Ribbon-to-BNC converter (same as hodoscopes)
while the other output is sent to a P/S Model 740 NIM Linear FI/FO summing modules. Each side of a layer is summed first (hA+, hA-, hB+, hB-, hC and hD sums). The sums are fed into
a LeCroy Model 428F summing module where layers hA+/- and hB+/- are summed to form hA and hB sums. A copy of each layer sum is sent to fADCs. The Pre-Shower SUM is then made from the sum of layer A, while the Shower SUM is made by summing all four layers. A copy of the Pre-Shower and Shower sums is also sent to fADCs. The Pre-Shower and Shower sums are then sent to a P/S Model 715
NIM Discriminator unit to form the PreShower Low/High (LO/HI) and Shower Low pre-triggers with thresholds \hPrShLo, \hPrShHi and \hSHLo, respectively with all gate widths set to 30 ns. A copy of the pre-triggers is sent to TDCs/Scaler modules for trigger timing and counting rate
information.

\subsection{Gas \v{C}erenkov Pre-Trigger} \label{hms_cer_section}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{./hCER_diagram.png}
  \caption{HMS Gas \v{C}erenkov Electronics Diagram. Same electronics diagram applies for SHMS Gas \v{C}erenkov. For original diagram, see Appendix \ref{appendix:Appx3} }
  \label{fig:hCER_diagram}
\end{figure}

@@

\newpage
\subsection{HMS Single Arm Pre-Trigger}\label{ssec:hms_single_arm_sec}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{HMS_SingleArm_diagram.png}
  \caption{HMS Single Arm Pre-Trigger Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx5}}
  \label{fig:HMS_SingleArm_diagram}
\end{figure}

@@

\subsection{Hodoscopes Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{pHODO_diagram.png}
  \caption{SHMS Hodoscopes Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx6}}
  \label{fig:pHODO_diagram}
\end{figure}
Each hodoscope plane consists of an array of scintillator bars coupled to a PMT at each end (See Figure \ref{fig:shms_stack}), so each bar reads out two signals. As shown in Figure \ref{fig:shms_hod_trg},
for example, hodoscope plane S1X consists of 26 signals (16 bars) read out in the Counting House (CH) patch. Each side of the plane (x13 signals/side) is fed into a 64-channel input passive splitter (16 Ch./set).
One-third (33\%) of the signal amplitude is sent via a 16-channel ribbon cable to a 64 Ch. input Ribbon-to-BNC converter (16 Ch./set) which outputs are fed into a 16-channel NIM input flash ADC (fADC).
The remaining two-thirds (66\%) of the signal amplitude is sent to a 16-Ch. input CAMAC Discriminator unit. The SHMS discriminators thresholds and gate widths were set to \shodthrs and \shodgate, respectively,
whereas the Quartz Discriminators thresholds and gate widths were set to \quartzthrs and \shodgate, respectively\\
\indent The discriminated signals are sent via two ribbon-cable outputs to CAEN1190 TDCs/Scalers (daisy-chained) and to a LeCroy 4564 CAMAC Logic unit to form the plane pre-triggers.
The Logic Unit takes four sets of 16-Ch. input ribbon cable and forms a 16-fold OR for each set by default. Further boolean operations are done through the module backplane by connecting a twisted pair cable to
the pin corresponding to the desired boolean operation. For hodoscope plane pre-triggers, the boolean operations are as follows:
\begin{empheq}[box=\fbox]{align*}
& \text{S1X = S1XL (13-fold OR) \textit{AND} S1XR (13-fold OR)} \\ 
& \text{S1Y = S1YT (14-fold OR) \textit{AND} S1YB (14-fold OR)} \\
& \text{S2X = S2XL (13-fold OR) \textit{AND} S2XR (13-fold OR)} \\ 
  & \text{S2Y = \big\{S2Y[1-16]T OR S2Y[17-21]T\big\} \textit{AND} \big\{S2Y[1-16]B OR S2Y[17-21]B\big\}}\footnotemark
\end{empheq}
\footnotetext{The SHMS S2Y (quartz) plane has 5 non-functional PMT channels on each side of the plane. In other words, it had up to 16 functional channels/side.}
\indent Once a pre-trigger has been made for each plane, they are sent to a NIM/ECL converter (Level Translator - Phillips Scientific (or P/S) Model 7126) via twisted pair cables to convert the ECL signal
(twisted pair) to a NIM signal. The NIM output is then sent to individual sets of a P/S Model 752 NIM Logic unit to adjust the widths of each of the plane pre-triggers as necessary before making a coincidence.
An X-Y hodoscope plane coincidence (S1 = S1X \textit{AND} S1Y, S2 = S2X \textit{AND} S2Y) is then made by feeding each hodoscope X-Y plane pair into a P/S Model 755 Nim Logic unit. A copy of each of the four
individual plane pre-triggers is also sent to another set of P/S Model 755 to make a 3/4 or 4/4 plane coincidence (via a front-panel knob) which defines the production hodoscope pre-trigger. A copy of all the
pre-triggers discussed above are sent to TDCs/Scalers via a NIM/ECL converter for timing and counting rate information. (See Figure \ref{fig:pHODO_diagram}) \\

\subsection{Pre-Shower / Shower Calorimeter Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{pCAL_diagram.png}
  \caption{SHMS PreShower / Shower Electronics Diagram. For original diagram, see Appendix \ref{appendix:Appx7}}
  \label{fig:pCAL_diagram}
\end{figure}

The SHMS Pre-Shower consists of two sets of fourteen (PMT-coupled) lead blocks oriented perpendicular to the Shower Counter blocks\cite{shms_preSh_talk}. The initial sum was done in the SHMS Electronics Hut. PMT signals
from groups of four blocks were summed to form:
\begin{empheq}[box=\fbox]{align*}
  & \text{preSh SUM [1-4]: [1-4]L + [1-4]R} \\ 
  & \text{preSh SUM [5-8]: [5-8]L + [5-8]R} \\ 
  & \text{preSh SUM [9-12]: [9-12]L + [9-12]R} \\ 
  & \text{preSh SUM [13-14]: [13-14]L + [13-14]R}
\end{empheq}
The Shower counter consists of 224 lead blocks, each coupled to a PMT at the end. Becasue of the high channel density of this detector, its signals are sent
directly to the ROC 4 fADCs in the SHMS detector hut.
\newpage
\subsection{Heavy/Noble Gas \v{C}erenkov Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{pCER_diagram.png}
  \caption{SHMS Gas \v{C}erenkovs Electronics Diagram.}
  \label{fig:pCER_diagram}
\end{figure}
\noindent The SHMS Heavy Gas \v{C}erenkov detector consists of a 1 m long, 1.6 m in diameter cylindrical tank located between the first set of hodoscope planes and the Aerogel (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four thin spherical mirrors that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_hgc_talk}. 
\newline
\indent The SHMS Noble Gas \v{C}erenkov detector consists of a 2 m long active length of Argon/Neon gas tank located before the first drift chamber (See Figure \ref{fig:shms_stack}).
The tank is filled with a gas and has four overlapping mirrors, that focus the \v{C}erenkov photons towards four 5-inch PMTs\cite{shms_ngc_talk}. 
For the original electronics diagram, See Appendix \ref{appendix:Appx4} on the HMS Gas \v{C}erenkov.
\subsection{Aerogel \v{C}erenkov Pre-Trigger}
\indent The SHMS Aerogel \v{C}erenkov detector consists of a 110 x 100 x 24.5 cm$^{3}$ rectangular aerogel tray coupled to a diffusion box. The diffusion box has seven 5-inch PMTs on each side
which detect \v{C}erenkov light produced by interactions with the Aerogel material\cite{shms_aero_article}. The detector is located between Heavy Gas \v{C}erenkov and second set of hodoscope planes
(See Figure \ref{fig:shms_stack}). The electronics diagram is the same as in Figure \ref{fig:pCER_diagram}. For the original diagram, see Appendix \ref{appendix:Appx4}
\newpage
\subsection{SHMS Single Arm Pre-Trigger}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{SHMS_SingleArm_diagram.png}
  \caption{SHMS Single Arm Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx8}.}
  \label{fig:SHMS_SingleArm_diagram}
\end{figure}

\indent The SHMS single arm trigger is be formed exactly as the HMS single arm trigger, with the exception of the detectors involved. Compare the electronics diagrams in Figures \ref{fig:SHMS_SingleArm_diagram} and
\ref{fig:HMS_SingleArm_diagram}, and read Section \ref{ssec:hms_single_arm_sec} for a detailed description of the electronic diagrams. 
\newpage
\section{Coincidence Trigger Set-Up}
\begin{figure}[h!]
  \centering
  \includegraphics[scale=0.35]{Coin_diagram.png}
  \caption{Coincidence Trigger Electronics Diagram. For the original diagram, see Appendix \ref{appendix:Appx9}}
  \label{fig:Coin_TRG}
\end{figure}
\indent In coincidence mode, the HMS and SHMS pre-triggers are sent to a NIM Logic module, where the first pre-trigger that arrives will open
a coincidence time window during which the second pre-trigger may or may not arrive in time. This will determine whether two events that occurred in
each spectrometer are correlated with the event that originated at the target. If the coincidence pre-trigger is formed, a copy will be sent to
scalers and TDCs, while the other copy is sent to a NIM/ECL converter where three copies (may be more) are sent via twisted pairs to every TDC module and the TI, which will be the \textbf{ONLY} Trigger Master in coincidence mode. Additional copies (not shown) may need to be sent to the HMS readout crate (ROC 01) as well, since the coincidence pre-trigger\footnote{A copy of the coincidence pre-trigger will also need to be sent to the
  HMS/SHMS TDC modules present in the detector huts crates via the patch panel.} is common to all crates. \\
\indent Once the coincidence pre-trigger is processed by the TI Master in ROC 02, it is sent to a NIM/ECL converter, where a copy is sent to scalers and TDCs, and the other copy is sent to the ``TRG'' NIM input in the TDC front panel, and daisy-chained with other TDCs in ROC02 and ROC01. The remaining crates in the HMS/SHMS huts receive a copy of the accepted trigger via fiber optics lines running from the TI Master to the TIs in all other crates. The TIs then distribute the accepted trigger to all modules of their respective crates.

\section{Electronic Dead Time Monitoring (EDTM)}\label{sec:edtm}
\noindent The EDTM system is a new method used in Hall C to measure the total dead time of the data acquisition (DAQ) system. It consists of introducing a controlled (fixed frequency) pulse
as near as possible to the detectors that form part of the trigger. Ideally, one would send the EDTM pulses at the detector level in the hut such that both the real physics and EDTM signals
pass through the same electronics. Since this is not easy or practical to do, the EDMT logic pulses are injected at the trigger logic level in the Counting Room.\\
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.54]{EDTM_diagram.png}
  \caption{EDTM electronics diagram.}
  \label{fig:EDTM_diagram}
\end{figure}
\indent By design, the EDTM is a \textit{real} trigger as measured by the electronics and readout systems. Since the EDTM is invasive to the trigger electronics, its frequency should be small
enough to minimize the probability of blocking actual physics triggers, but sufficiently large to gather the necessary statistics for a precise \textit{dead time} measurement during the course of a run. \\
\indent Figure \ref{fig:EDTM_diagram} shows a simplified diagram of the EDTM signal distribution through the trigger electronics. The EDTM logic signals (purple) are injected into the trigger logic
where they mix with the physics pre-triggers (magenta). A separate copy of the EDTM is also sent to scalers/TDCs to be used in the dead time calculation. If the
EDTM makes it to the front-end of the Trigger Interface (TI) module and gets accepted (L1 Accept), it has esentially measured both the electronics and computer dead time.\\
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.40]{EDTM_TI.png}
  \caption{Cartoon representation of EDTM (purple) and physics (magenta) pre-triggers at the TI module front-end.}
  \label{fig:EDTM_TI}
\end{figure}\\
\indent Figure \ref{fig:EDTM_TI} shows the random physics (magenta) and clocked (purple) EDTM pulses at an input channel of the TI front-end where the EDTM has
been set to a sufficiently large frequency ($1/T_{\mathrm{clk}}$) to ensure that enough EDTM signals get accepted in order to make a statistically significant and
reliable dead time calculation. In this example, an EDTM signal has been accepted by the TI which triggered a \textit{BUSY} signal for a time $\tau$ during which all other incoming pre-triggers are blocked
contributing to the DAQ computer dead time. The accepted pre-triggers are distributed to all ROCs for data readout. \\
\indent Over the course of a run, the total dead time ($T_{\mathrm{TDT}}$), or alternatively, the total live time ($T_{\mathrm{TLT}}$) in terms of the EDTM is defined as
\begin{equation}
 T_{\mathrm{TDT}} \equiv 1 - T_{\mathrm{TLT}} = 1 - \frac{N_{\mathrm{edtm,acc}}}{N_{\mathrm{edtm,scl}}},
  \label{eq:3.24}
\end{equation}
where $N_{\mathrm{edtm,acc}}$ is the number of accepted EDTM counts obtained by requiring a non-zero hit on the EDTM TDC spectrum, and $N_{\mathrm{edtm,scl}}$ is the number of EDTM scaler counts regardless of whether or not
the EDTM was accepted. In reality, frequent beam trips occur during the course of a run which makes this calculation biased since one can measure live times of $\sim$100$\%$ during beam-off periods as only
the EDTM signal (and cosmic rays) are measured. To eliminate this bias, the live time calculation was done by making a software cut on the beam current above a certain threshold.
Furthermore, since the EDTM events are generated by a clock, rather than a poisson source, it introduces an additional bias since the EDTM cannot block itself. To account for this bias, an additional correction
to the total live time was derived and can be found in Ref.\cite{DMack_livetime_2019}. This correction, however, is negligible provided that the EDTM rate is sufficiently low as was the case for this experiment ($\sim 2$ Hz). \\
\indent Even though only the total live time is required as a correction factor in the measured cross section, one may also calculate the computer live time defined as
\begin{equation}
  T_{\mathrm{CLT}} = \frac{N_{\mathrm{phy,acc}}}{N_{\mathrm{phy,scl}}},
  \label{eq:3.25}
\end{equation}
where $N_{\mathrm{phy,acc}}$ is the number of accepted physics triggers obtained by requiring a zero hit on the EDTM TDC spectrum (EDTM rejected by TI) and $N_{\mathrm{phy,scl}}$ is the number of physics trigger scaler counts after
having subtracted the EDTM scaler counts. The electronic live time can then be obtained from the following formula:
\begin{equation}
  T_{\mathrm{TLT}} = T_{\mathrm{CLT}} \cdot T_{\mathrm{ELT}},
  \label{eq:3.26}
\end{equation}
